---
title: "synthetic-synthetic_genome-analysis"
author: "Francine Camacho"
date: "7/23/2019"
output:
  html_document: 
    smart: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/HMM-analysis/')
```

```{r, message = FALSE }
# load neccessary packages 
require(tidyverse) 
require(ggsci)
require(ggpubr)
```

### Load segmented profiled siderophore HMMs for synthetic genomes
```{r, message = FALSE}
#load siderophore HMM data and recode sampleType for the complexity of the synthetic sample (#of genomes in samples)
siderophore <- read_delim("combined-Siderophores-spHMM-synthetic_genomes-results.txt",col_names = F, delim = "\t") 
names(siderophore) <- c("readID", "sampleType", "sampleID", "cyclaseType", "HMMScore", "window","interval")

siderophore$sampleType[str_detect(siderophore$sampleID, "high") ==TRUE] <-"high"
siderophore$sampleType[str_detect(siderophore$sampleID, "low") ==TRUE] <-"low"

siderophore$interval <- factor(siderophore$interval,levels = c(
"0_30","10_40","20_50","30_60","40_70","50_80","60_90","70_100","80_110",
"90_120","100_130","110_140","120_150","130_160","140_170","150_180", "160_190",
"170_200", "180_210","190_220", "200_230", "210_240", "220_250", "230_260",
"240_270", "250_280", "260_290", "270_300", "280_310", "290_320", "300_330",
"310_340", "320_350", "330_360", "340_370", "350_380", "360_390", "370_400",
"380_410", "390_420", "400_430", "410_440", "420_450", "430_460", "440_470",
"450_480"))

# because of gaps in the HMM we will ignore the following intervals 
ignored_intervals  <- c("50_80","60_90","70_100","80_110",
"90_120","100_130","110_140","120_150","130_160","140_170","150_180", "160_190",
"170_200", "180_210","190_220", "200_230", "210_240", "220_250", "230_260","410_440", "420_450")

siderophore_df <- siderophore %>% filter(!interval %in% ignored_intervals )
```

##### Remove the reading frame information from the readID. 
```{r,message=FALSE}
# Function to aggregate identitical sample reads located at different reading frames 
# and take the frame with the highest HMM score
formatHMM<-function(hmmdf){
  hmmdfRecoded <- separate(hmmdf, readID, into = c("readIDOnly","F_R_read_frame"), sep = "/", extra = "merge")
  hmmdfRecoded_FR <- separate(hmmdfRecoded, F_R_read_frame, into = c("F_R","frameNumb"), sep = "_", extra = "merge")
  hmmdfRecodedDF<- within(hmmdfRecoded_FR, readID <- paste(readIDOnly,F_R, sep='/'))
  hmmdfRecodedDFUnique<-aggregate(HMMScore ~ readID + sampleID + sampleType + cyclaseType + window + interval, hmmdfRecodedDF, max)
  colnames(hmmdfRecodedDFUnique)<-c("readID","Sample", "sampleType", "cyclaseType", "window", "interval","HMMScore")
  return(hmmdfRecodedDFUnique)
}

#Keep duplicated reads if they are in different reads
siderophore_recoded<-formatHMM(siderophore_df)
```

##### Load the BLAST data for Siderophore genes against our synthetic dataset. 
```{r, message=FALSE}
# BLAST unfiltered reads at 95% pident no readCoverage filter
all_sidBlastDF <- read_delim("../Siderophores-TP-reads/sidBlastDF_all-results.txt", col_names = T, delim = "\t")

```

##### Positional information about siderophore domains and their locations in respect to the siderophore 30_10 spHMM models 
```{r, message=FALSE}
#load siderophore interval positions data 
sid_positions <- read_tsv("../Siderophores-TP-reads/siderophore_gene-interval_position-data-updated.txt",col_names = T)
```

#### Positional read analysis in respect to location mapped to siderophore domain 
##### Keep reads that map to the interval of a given model and covers the model 90% 
```{r, message=FALSE}
filter_blast <- function(df, pos_df){
  results <- data.frame()
  for (i in 1:nrow(pos_df)){
    gene_data <- pos_df[i,]
    print(gene_data$gene_name)
    #filters datatframe for edges and internal reads compared to model interval
    interval_df_1 <- df %>% filter(qseqid ==gene_data$gene_name) %>%
      filter(qstart %in% gene_data$start:gene_data$end | qend %in% gene_data$start:gene_data$end)
    # need to get reads that are bigger than the interval 
    interval_df_2 <- df %>% filter(qseqid ==gene_data$gene_name) %>% 
      filter(qstart < gene_data$start & qend > gene_data$end)
    interval_df <- rbind(interval_df_1,interval_df_2)
    if (nrow(interval_df) > 0){
      for (j in 1:nrow(interval_df)){
        in_interval_count <- sum(interval_df[j,]$qstart:interval_df[j,]$qend %in% gene_data$start:gene_data$end)
        interval_count <- length(gene_data$start:gene_data$end)
        interval_cov <- (in_interval_count/interval_count) * 100
        if (interval_cov >=90){
          res_df<-interval_df[j,]
          res_df$model_cov <- interval_cov
          res_df$interval <- gene_data$interval
          results <- rbind(results,res_df)
        }
      }
      
    }
    
  }
  return(results)
}
```

```{r, message=FALSE}
#Filter BLAST reads that are within the Sid genes intervals and cover 90% of the model interval
#sid_blast_intervals<- filter_blast(all_sidBlastDF, sid_positions)
sid_blast_intervals <- read_tsv("../Siderophores-TP-reads/sid_blast_intervals-updated.txt", col_names = T)
```

##### Determine manual cutoffs for each interval within the model

```{r, message=FALSE}
compare_reads <- function(hmm_df, blast_df){
  names(blast_df)[names(blast_df)=="sseqid"] <-"readID"
  names(blast_df)[names(blast_df)=="qseqid"] <-"bgcName"
  names(hmm_df)[names(hmm_df)=="cyclaseType"] <-"cyclase_type"
  
  # remove columns to compare the two dataframe 
  blastDF <- blast_df %>% select(-c(model_cov, interval))
  common_reads <- hmm_df %>% semi_join(.,blastDF) 
  common_reads$readCheck<-"common-read"
  hmm_unique_reads <- hmm_df %>% anti_join(.,blastDF)
  hmm_unique_reads$readCheck<-"hmm-unique-read"
  compared_data <- rbind(common_reads, hmm_unique_reads)
  return(compared_data)
}

sid_bin <- compare_reads(siderophore_recoded, sid_blast_intervals)
names(sid_bin)[names(sid_bin)=="cyclase_type"] <-"cyclaseType"
```

##### Trying to determine the HMM cutoffs; calculate median 
```{r, message=FALSE}
sid_bin %>% filter(readCheck == "common-read") %>% group_by(interval) %>% mutate(medianScore = round(median(HMMScore))) %>% distinct(interval, readCheck, medianScore) %>% ungroup()
```

##### Remove HMM unique reads that mapped to BLAST but didn't make the 90% interval model coverage cutoff imposed. These reads shouldn't be considered hmm-unique
```{r, message=FALSE}
#Function to find the hmm_unique reads that are also mapped using BLAST but didn't make 90% interval
compare_hmm_unique <- function(hmm_df, blast_df, pos_df){

  #hmm_unique_df <- hmm_df  %>% inner_join(.,blast_df, by = c("readID"="sseqid", "Sample", "sampleType","cyclaseType"= "cyclase_type"))
   hmm_unique_df <- hmm_df  %>% inner_join(.,blast_df, by = c("readID"="sseqid", "Sample", "sampleType", "cyclase_type"))
  intervals <- unique(hmm_unique_df$interval)
  results <- data.frame()
  #check that reads are in the same interval to throw out 
  for (i in 1:length(intervals)){
    gene_interval_data <- pos_df %>% filter(interval == intervals[i])
    for (k in 1:nrow(gene_interval_data)){
        gene_data <- gene_interval_data[k,]
        #filters datatframe for edges and internal reads compared to model interval
        interval_df  <- hmm_unique_df %>% filter(qseqid ==gene_data$gene_name) %>%
          filter(qstart %in% gene_data$start:gene_data$end | qend %in% gene_data$start:gene_data$end)
        if (nrow(interval_df) > 0 ){
           results <- rbind(results,interval_df)
        }
    
  }
 }  
  return(results%>% distinct())
}
```

### Apply different HMM Scores cutoff to the HMM dataset
```{r,message=FALSE}
#Filter siderophore data with cutoffs to compare to BLAST interval reads 
sid_filtered_median <- siderophore_recoded %>% filter((interval == "0_30" & HMMScore>=30)|
                                         (interval == "10_40" & HMMScore >=35)| 
                                         (interval == "20_50" & HMMScore >=30)| 
                                         (interval == "30_60" & HMMScore >=30)| 
                                         (interval == "40_70" & HMMScore >=20 )|
                                         (interval == "240_270" & HMMScore>=25)| 
                                         (interval == "250_280" & HMMScore >=25 )| 
                                         (interval == "260_290" & HMMScore >=25 )| 
                                         (interval == "270_300" & HMMScore >=20 )| 
                                         (interval == "280_310" & HMMScore >=20 )| 
                                         (interval == "290_320" & HMMScore >=20 )| 
                                         (interval == "300_330" & HMMScore >=25 )| 
                                         (interval == "310_340" & HMMScore >=25 )| 
                                         (interval == "320_350" & HMMScore >=25 )| 
                                         (interval == "330_360" & HMMScore >=25 )| 
                                         (interval == "340_370" & HMMScore >=30 )| 
                                         (interval == "350_380" & HMMScore >=30 )| 
                                         (interval == "360_390" & HMMScore >=25 )|
                                         (interval == "370_400" & HMMScore >=20 )| 
                                         (interval == "380_410" & HMMScore >=25 )| 
                                         (interval == "390_420" & HMMScore >=25 )| 
                                         (interval == "400_430" & HMMScore >=20 )|
                                         (interval == "430_460" & HMMScore >=20 )| 
                                         (interval == "440_470" & HMMScore >=20 )|
                                         (interval == "450_480" & HMMScore >=20 ))

sid_filtered_median_subfive <- siderophore_recoded %>% filter((interval == "0_30" & HMMScore>=25)|
                                         (interval == "10_40" & HMMScore >=30)| 
                                         (interval == "20_50" & HMMScore >=25)| 
                                         (interval == "30_60" & HMMScore >=25)| 
                                         (interval == "40_70" & HMMScore >=15 )|
                                         (interval == "240_270" & HMMScore>=20)| 
                                         (interval == "250_280" & HMMScore >=20 )| 
                                         (interval == "260_290" & HMMScore >=20 )| 
                                         (interval == "270_300" & HMMScore >=15 )| 
                                         (interval == "280_310" & HMMScore >=15 )| 
                                         (interval == "290_320" & HMMScore >=15 )| 
                                         (interval == "300_330" & HMMScore >=20 )| 
                                         (interval == "310_340" & HMMScore >=20 )| 
                                         (interval == "320_350" & HMMScore >=20 )| 
                                         (interval == "330_360" & HMMScore >=20 )| 
                                         (interval == "340_370" & HMMScore >=25 )| 
                                         (interval == "350_380" & HMMScore >=25 )| 
                                         (interval == "360_390" & HMMScore >=20 )|
                                         (interval == "370_400" & HMMScore >=15 )| 
                                         (interval == "380_410" & HMMScore >=20 )| 
                                         (interval == "390_420" & HMMScore >=20 )| 
                                         (interval == "400_430" & HMMScore >=15 )|
                                         (interval == "430_460" & HMMScore >=15 )| 
                                         (interval == "440_470" & HMMScore >=15 )|
                                         (interval == "450_480" & HMMScore >=15 ))

sid_filtered_median_plusfive <- siderophore_recoded %>% filter((interval == "0_30" & HMMScore>=35)|
                                         (interval == "10_40" & HMMScore >=40)| 
                                         (interval == "20_50" & HMMScore >=35)| 
                                         (interval == "30_60" & HMMScore >=35)| 
                                         (interval == "40_70" & HMMScore >=25 )|
                                         (interval == "240_270" & HMMScore>=30)| 
                                         (interval == "250_280" & HMMScore >=30 )| 
                                         (interval == "260_290" & HMMScore >=30 )| 
                                         (interval == "270_300" & HMMScore >=25 )| 
                                         (interval == "280_310" & HMMScore >=25 )| 
                                         (interval == "290_320" & HMMScore >=25 )| 
                                         (interval == "300_330" & HMMScore >=30 )| 
                                         (interval == "310_340" & HMMScore >=30 )| 
                                         (interval == "320_350" & HMMScore >=30 )| 
                                         (interval == "330_360" & HMMScore >=30 )| 
                                         (interval == "340_370" & HMMScore >=35 )| 
                                         (interval == "350_380" & HMMScore >=35 )| 
                                         (interval == "360_390" & HMMScore >=30 )|
                                         (interval == "370_400" & HMMScore >=25 )| 
                                         (interval == "380_410" & HMMScore >=30 )| 
                                         (interval == "390_420" & HMMScore >=30 )| 
                                         (interval == "400_430" & HMMScore >=25 )|
                                         (interval == "430_460" & HMMScore >=25 )| 
                                         (interval == "440_470" & HMMScore >=25 )|
                                         (interval == "450_480" & HMMScore >=25 ))


```

```{r, message=FALSE}
#return the hmm-unique reads 
return_hmm_unique <- function(hmm_df, blast_df){
  oxyn_hmm_df <- hmm_df  %>% select(-c(window))
  oxyn_hmm_df$interval <- as.character(oxyn_hmm_df$interval)
  names(oxyn_hmm_df)[names(oxyn_hmm_df)=="cyclaseType"] <-"cyclase_type"
  names(blast_df)[names(blast_df)=="sseqid"] <-"readID"
  OxyN_hmm_unique <- oxyn_hmm_df %>% anti_join(.,blast_df, by= c("readID", "Sample", "sampleType", "interval"))
  return(OxyN_hmm_unique)
}

sid_positions$cyclase_type <-"siderophore"
all_sidBlastDF$cyclase_type <-"siderophore"
median_sid_hmmunique <-return_hmm_unique(sid_filtered_median, sid_blast_intervals)

median_sid_hmmunique_less_model_cov<-compare_hmm_unique(median_sid_hmmunique,all_sidBlastDF,sid_positions ) 

median_remaining_hmm_sid<- median_sid_hmmunique %>% anti_join(.,median_sid_hmmunique_less_model_cov)
table(median_remaining_hmm_sid$interval)

sid_genomes_median_check <- c("gi|529222834|ref|NC_021985.1|", "CP011998.1", "CP012944.1", "DQ381420.1", "gi|640935821|gb|JJOH01000007.1|", "AZXI01000015.1", "JOFH01000011.1", "gi|928416322|gb|LGKI01000367.1|", "CP014065.1", "gi|1081599680|emb|FMWJ01000019.1|","HE965803.1", "CP016022.1", "JOFH01000020.1")


#median_remaining_hmm_sid %>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% distinct(scaffold_id)

remaining_hmm_sid_pos_median_check <- median_remaining_hmm_sid %>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% filter(scaffold_id %in% sid_genomes_median_check)
############################################################################################################################
subfive_sid_hmmunique <-return_hmm_unique(sid_filtered_median_subfive, sid_blast_intervals)

subfive_sid_hmmunique_less_model_cov<-compare_hmm_unique(subfive_sid_hmmunique,all_sidBlastDF,sid_positions ) 

subfive_remaining_hmm_sid<- subfive_sid_hmmunique %>% anti_join(.,subfive_sid_hmmunique_less_model_cov)
table(subfive_remaining_hmm_sid$interval)

#subfive_remaining_hmm_sid %>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% distinct(scaffold_id)
sid_genomes_subfive_check <- c("CP011998.1", "gi|640935821|gb|JJOH01000007.1|", "CP012944.1", "gi|529222834|ref|NC_021985.1|", "CP014065.1","gi|1081599680|emb|FMWJ01000019.1|", "DQ381420.1", "AZXI01000015.1", "JOFH01000011.1", "gi|928416322|gb|LGKI01000367.1|","JOFH01000020.1", "HE965803.1", "CP016022.1", "JOFH01000009.1")

remaining_hmm_sid_pos_subfive_check <- subfive_remaining_hmm_sid %>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% filter(scaffold_id %in% sid_genomes_subfive_check)

###############################################################################################################################################

plusfive_sid_hmmunique <-return_hmm_unique(sid_filtered_median_plusfive, sid_blast_intervals)

plusfive_sid_hmmunique_less_model_cov<-compare_hmm_unique(plusfive_sid_hmmunique,all_sidBlastDF,sid_positions ) 

plusfive_remaining_hmm_sid<- plusfive_sid_hmmunique %>% anti_join(.,plusfive_sid_hmmunique_less_model_cov)
table(plusfive_remaining_hmm_sid$interval)

#plusfive_remaining_hmm_sid %>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% distinct(scaffold_id)

sid_genomes_plusfive_check <- c("gi|529222834|ref|NC_021985.1|", "JOFH01000011.1", "AZXI01000015.1", "CP012944.1",
                                "gi|928416322|gb|LGKI01000367.1|", "CP014065.1", "CP011998.1", "HE965803.1")

remaining_hmm_sid_pos_plusfive_check <- plusfive_remaining_hmm_sid %>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% filter(scaffold_id %in% sid_genomes_plusfive_check)

```

```{r}
#Function to parse out HMM unique reads for each sample. 
parseReads <- function(HMMdf, modelName, dirname){
  setwd(dirname)
  samples<-unique(HMMdf$Sample)
  for (s in 1:length(samples)){
    currentSample<-samples[s]
    currentSampleResults<- HMMdf %>% filter(Sample == currentSample)
    currentSampleReads<- unique(currentSampleResults$readID)
    fileName<-paste0(currentSample,paste("-",modelName,"-hmm_unique",sep =""), ".txt", sep ="")
    write.table(currentSampleReads,fileName, quote = F, row.names = F, col.names = F )
  }
}

#parseReads(remaining_hmm_sid_pos_median_check, "siderophore", "/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/HMM-analysis/hmm-unique-analysis/median_score/")

#parseReads(remaining_hmm_sid_pos_subfive_check, "siderophore", "/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/HMM-analysis/hmm-unique-analysis/minus_five_score/")

#parseReads(remaining_hmm_sid_pos_plusfive_check, "siderophore", "/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/HMM-analysis/hmm-unique-analysis/plus_five_score/")

```

```{r}

blast_hmm_unique <- read_tsv("/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/HMM-analysis/hmm-unique-analysis/sid_hmm-unique_blast-results.txt", col_names = T)

filter_hmm_unique  <- function(df, pos_df){
  results <- data.frame()
  for (i in 1:nrow(df)){
    blast_data <- df[i,]
    domain <- blast_data$sseqid
    domain_interval <- blast_data$interval 
    read_start <-blast_data$sstart
    read_end <-blast_data$send
    min_start <- min(read_start, read_end)
    max_end <- max(read_start, read_end)
    gene_data <- pos_df %>% filter( gene_name == domain  & interval == domain_interval)
    #filters dataframe for edges and internal reads compared to model interval
    interval_df_1 <- blast_data %>% filter(sseqid ==gene_data$gene_name) %>%
      filter(min_start %in% gene_data$start:gene_data$end | max_end %in% gene_data$start:gene_data$end)
    # need to get reads that are bigger than the interval 
    interval_df_2 <- df %>% filter(sseqid ==gene_data$gene_name) %>% 
      filter(min_start < gene_data$start & max_end > gene_data$end)
    
    results <- rbind(results,interval_df_1,interval_df_2)
  }
  return(results)
}

median_blast_df <- blast_hmm_unique %>% filter(cutoff== "median_score") %>% inner_join(., remaining_hmm_sid_pos_median_check, by = c("qseqid" = "readID", "Sample"))
subfive_blast_df <- blast_hmm_unique %>% filter(cutoff== "minus_five_score") %>% inner_join(., remaining_hmm_sid_pos_subfive_check, by = c("qseqid" = "readID", "Sample"))

plusfive_blast_df <- blast_hmm_unique %>% filter(cutoff== "plus_five_score") %>% inner_join(., remaining_hmm_sid_pos_plusfive_check, by = c("qseqid" = "readID", "Sample"))


remove_sid_reads_median <- filter_hmm_unique(median_blast_df,sid_positions) %>% as.data.frame() %>% distinct(qseqid,Sample)  
names(remove_sid_reads_median)[1]<-"readID"
remove_sid_reads_subfive <- filter_hmm_unique(subfive_blast_df,sid_positions) %>% as.data.frame() %>% distinct(qseqid, Sample) 
names(remove_sid_reads_subfive)[1]<-"readID"
remove_sid_reads_plusfive <- filter_hmm_unique(plusfive_blast_df,sid_positions) %>% as.data.frame()%>% distinct(qseqid, Sample)
names(remove_sid_reads_plusfive)[1]<-"readID"
```

```{r}
calculate_F1 <- function(hmm_df, hmm_fp, blast_df, intervals){
  #added this because factor vector 
  hmm_df$interval <- as.character(hmm_df$interval)
  hmm_df <- hmm_df %>% select(-c(window))
  names(hmm_df)[names(hmm_df)=="cyclaseType"] <-"cyclase_type"
  names(blast_df)[names(blast_df)=="sseqid"] <-"readID"
  
  results<-data.frame(interval=character(), F1=numeric())
  for (i in 1:length(intervals)){
    model <- intervals[i]
    model_hmm <- hmm_df %>% filter(interval== model) # filter out model HMM results 
    model_blast <- blast_df %>% filter(interval== model) 
    model_hmm_fp <-hmm_fp %>% filter(interval== model) 
    TP <- model_hmm %>% inner_join(.,model_blast) %>% nrow()
    FP <- model_hmm_fp %>% nrow()
    FN <- model_blast %>% anti_join(.,model_hmm) %>% nrow()
    F1_metric <- (2*TP)/((2*TP)+FP+FN)
    row <- data.frame(cbind(model, F1_metric))
    results<- results %>% add_row(interval = model, F1 = F1_metric)
  }
  

    return(results)
  
}

# Genome gi|928417145|gb|LGKI01000307.1| is a partial contig and TPs therefore for the F1 analysis we will ignore the genome from HMM data, false positives, and BLAST data 
subfive_false_postives <- subfive_remaining_hmm_sid %>% anti_join(.,remove_sid_reads_subfive)  %>% filter(!str_detect("LGKI01000307.1",readID))


median_false_postives <- median_remaining_hmm_sid %>% anti_join(.,remove_sid_reads_median)  %>% filter(!str_detect("LGKI01000307.1",readID))


plusfive_false_postives <- plusfive_remaining_hmm_sid %>% anti_join(.,remove_sid_reads_plusfive)  %>% filter(!str_detect("LGKI01000307.1",readID))


sid_f1_cutoff_median<- calculate_F1(sid_filtered_median %>% filter(!str_detect("LGKI01000307.1",readID)), median_false_postives, sid_blast_intervals %>% filter(!str_detect("LGKI01000307.1",sseqid)), unique(siderophore_recoded$interval))
sid_f1_cutoff_median$cutoff<- "median"
sid_f1_cutoff_plusfive<- calculate_F1(sid_filtered_median_plusfive %>% filter(!str_detect("LGKI01000307.1",readID)), plusfive_false_postives, sid_blast_intervals %>% filter(!str_detect("LGKI01000307.1",sseqid)), unique(siderophore_recoded$interval))
sid_f1_cutoff_plusfive$cutoff<- "+5"
sid_f1_cutoff_subfive<- calculate_F1(sid_filtered_median_subfive %>% filter(!str_detect("LGKI01000307.1",readID)), subfive_false_postives, sid_blast_intervals %>% filter(!str_detect("LGKI01000307.1",sseqid)), unique(siderophore_recoded$interval))
sid_f1_cutoff_subfive$cutoff<- "-5"

sid_f1_df <- rbind(sid_f1_cutoff_median,sid_f1_cutoff_plusfive,sid_f1_cutoff_subfive)

```

```{r}
supp_fig_sid_f1 <-ggplot(data = sid_f1_df, mapping = aes(x = interval, y = F1, group= cutoff, colour=cutoff )) + geom_point()+
    geom_line() + ylim(0,1)+ geom_hline(yintercept=c(.50,0.2151334), linetype="dashed", 
                color = c("red", "green"), size=1) + theme_pubclean() +  scale_color_npg(name="HMM Score Cutoff")

ggsave(supp_fig_sid_f1, file="siderophore_synthetic_F1-final.eps", device="eps", width = 17, height = 7)
```
#Find the optimal cutoffs per interval using F1 metric
```{r}
sid_f1_df %>% group_by(interval) %>% top_n(1, F1) %>% ungroup() %>%  filter(F1 >=0.5) %>% arrange(interval) %>% as.data.frame()
```

## Make final siderophore with final cutoffs for the highest F1 for each interval 
```{r}
final_ggplot_sid<- sid_f1_df %>% group_by(interval) %>% top_n(1, F1) %>% ungroup() %>% ggplot(., mapping = aes(x = interval, y = F1, group= 1 )) + geom_point()+  ylim(0,1)+
    geom_line() +  geom_hline(yintercept=.50, linetype="dashed", 
                color = "red", size=1) + theme_pubclean()+ scale_color_npg()


#ggsave(final_ggplot_sid, file="siderophore_synthetic_F1-final.eps", device="eps", width = 17, height = 7)
```


```{r}
# Finalized cutoffs with intervals and cutoffs 
sid_filtered_final <- siderophore_recoded %>% filter(!str_detect("LGKI01000307.1",readID)) %>% filter((interval == "0_30" & HMMScore>=25)|
                                         (interval == "10_40" & HMMScore>=30)| 
                                         (interval == "20_50" & HMMScore >=25 ) | 
                                         (interval == "30_60" & HMMScore >=25 )| 
                                         (interval == "240_270" & HMMScore >=20 )| 
                                         (interval == "250_280" & HMMScore >=25 )|
                                         (interval == "260_290" & HMMScore >=25 )|
                                         (interval == "300_330" & HMMScore >=20 )|
                                         (interval == "320_350" & HMMScore >=20 )|
                                         (interval == "330_360" & HMMScore >=20 )|
                                         (interval == "340_370" & HMMScore>=25)|
                                         (interval == "350_380" & HMMScore>=25)|
                                         (interval == "360_390" & HMMScore>=20))


#sid_filtered_final%>% separate(readID, c("scaffold_id", "rest_id"), sep = "-", remove = F) %>% distinct(scaffold_id)
write_tsv( sid_filtered_final %>% select(-c(window, interval))%>% group_by(readID,Sample) %>% arrange(desc(HMMScore)) %>% top_n(1, HMMScore) %>% filter(row_number() == 1) %>% ungroup(), "siderophore-final_data-raw.txt", col_names = T)

#rename siderophore hmm_unique within the folder to detected_reads
parseReads(sid_filtered_final, "siderophore", "/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/read-quantification/sample-readIDs-files")
#parseReads(subfive_false_postives, "siderophore-fps", "/Users/francinecamacho/Google Drive/T2PKS-paper/computational_analysis/synthetic_genomes/revision-synthetic-genomes/other-BGCs-poc/siderophores/HMM-analysis/false-positives-subfive/")
```
## Output FP/TP for HMM reads detected in final models with final cutoffs/ keep reads duplicated in multiple intervals with the highest HMM Score if, they are tied, then choose the first one. 
```{r}

sid_filtered_final_distinct <- sid_filtered_final %>% arrange(desc(HMMScore)) %>% group_by(readID, Sample) %>% top_n(1) %>% 
  filter(row_number()==1) %>% ungroup() %>% select(-c(cyclaseType))

sid_filtered_final_distinct$interval <-as.character(sid_filtered_final_distinct$interval)
#names(sid_filtered_final_distinct)[names(sid_filtered_final_distinct)=="cyclaseType"] <-"cyclase_type"

results_common<-sid_filtered_final_distinct %>% inner_join(., sid_blast_intervals %>% filter(!str_detect("LGKI01000307.1",sseqid)), by = c("readID" = "sseqid", "Sample", "sampleType", "interval")) %>% select(-c(model_cov))
results_common$cyclase_type <- "siderophore"

results_hmm_unique<-sid_filtered_final_distinct %>% anti_join(., sid_blast_intervals %>% filter(!str_detect("LGKI01000307.1",sseqid)), by = c("readID" = "sseqid", "Sample", "sampleType", "interval")) 
results_hmm_unique$cyclase_type <- "siderophore"

results_hmm_unique_TP<-compare_hmm_unique(results_hmm_unique,all_sidBlastDF,sid_positions ) 


results_hmm_unique_FP <- results_hmm_unique %>% anti_join(.,results_hmm_unique_TP)


sid_filtered_final_distinct_TPs <- rbind(results_common, results_hmm_unique_TP) %>% select(-c(7:9,11:15))

combined_sid_results <- sid_filtered_final_distinct_TPs %>% full_join(., results_hmm_unique_FP )

write_tsv(combined_sid_results,"sid-final_models-data.txt",col_names = T)
```

```{r}
sessionInfo()
```